AtmoSwing (v1.4): Analog Technique Model for Statistical weather forecastING

### 2015-11-02 AtmoSwing - Geoscientific Model Development/template.tex
Corrections from Michel.

When using a parameterization of the AM, previously optimized in perfect prognosis, in operational forecasting, we should take into account the increasing uncertainty of the predictor with the lead time. An option, proposed by \citet{Thevenot2004}, is to increase the number of analogues with the lead time. These values should be optimized, for every lead time, by using reforecast data of the model considered for operational forecasting.

@@

Another reason for working in a perfect prognosis framework is that numerical models evolve continuously, and so does the forecast they provide. Re-forecasts would allow us to work on a homogeneous dataset, as they are regularly reprocessed. However, we would need to redo the calibration procedure every time a new version is available, in order to reduce the bias \citep{Wilson2002}. Moreover, the reforecasts are not re-processed for every new version of the model, meaning we still end up with a bias between the forecast and the archive. Finally, the length of reanalyses datasets are usually much longer than reforecast datasets, which allows us to identify more robust relationships. The size of the archive is indeed an important criteria for the AM.

@@

We mainly rely on the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows assessing the predicted cumulative distribution functions $F(y)$ compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

\begin{equation}
\label{eq:CRPSS}
CRPSS = \dfrac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\dfrac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference and $CRPS_{p}$ would be the one for a perfect forecast ($CRPS_{p} = 0$).

@@

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criteria) over a large domain. Every point (or cell) of the full domain is assessed jointly on all predictors of the level of analogy (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The detailed stages are the followings: (i) The unitary spatial window is expanded in every 4 directions successively. The performance score is processed for these 4 windows. (ii) Only the direction providing the best improvement is applied to our spatial window. (iii) From this new spatial window, an increase in every 4 directions is once again assessed, and the best improvement is applied. (iv) The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analogues is optimized for the current level of analogy.
		\item A new level of analogy can be added, based on other variables (such as the moisture index) on predefined pressure levels and time frames. The number of analogues for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (a) for the new level. The parameters calibrated on the previous analogue levels are fixed and do not change (except the number of analogues, at the final stage). 
	\end{enumerate}
	\item Finally, the numbers of analogues are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogues of each level in a systematic way.
\end{enumerate}

@@

There are numerous implementation variants of GAs often optimal for a given problem \citep{Hart1991a,Schraudolph1992a}. However, the structure of the method (Figure \ref{fig:structure_gas}) resulting from the work of \citet{Holland1992b} is common to most applications \citep{Back1993b}. The divergences are the operators implementation, through significantly different algorithms, which has an important effect on the results \citep{Gaffney2010a}.

@@

	\item \textbf{Tournament selection} \citep{Michalewicz1996, Zitzler2004a}: two individuals are randomly selected and fight. The one with the highest score is chosen, but with a certain probability, in order to reduce the selection pressure. This procedure is repeated until the mating pool is full. Individuals can be selected several times, and thus be represented several times in the mating pool.
\end{itemize}

@@

	\textit{Roulette wheel weighting on fitness}: the selection probability is calculated based on the value of the objective function. This approach gives more weight to the best individuals when the distribution of scores is wide, while the weight is almost similar when all individuals have approximately the same score \citep{Haupt2004}. The probability $p_{n}$ of each individual is calculated by the equation \ref{equation_mating_score_weighting}:
	\begin{equation}
	p_{n}=\frac{score_{n}-score_{N_{mp}}}{\sum_{n=1}^{N_{mp}} (score_{n}-score_{N_{mp}})}
	\label{equation_mating_score_weighting}
	\end{equation}
	In our application, the last individual ($N_{mp}$) has zero probability of being selected.

@@

	\item \textbf{Individual adaptive search radius}: based on the ideas of the non-uniform mutation, we introduce a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1, is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is considered initially as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_rayon_adaptatif}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, in the same way as for the non-uniform mutation. No external parameter is therefore necessary.

@@

	\item \textbf{Chromosome of adaptive search radius}: we introduced this operator that combines the operations of the chromosome of adaptive mutation rate to our adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second containing the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameters are required.

	\item \textbf{Multi-scale mutation}: finally, we developed another approach, that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea is consistent as long as we are confident that the algorithm will converge towards the global optimum. Indeed, once the algorithm is in the exploitation mode, it is very unlikely to go out of the region it converges to. We wanted to test an approach that combines both exploration and exploitation during the whole optimization. Thus, we considered the search radius $r_{a}$ of equation \ref{equation_mutation_rayon_adaptatif} as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02. The only external parameter is the mutation rate which is fixed.

@@

We used a process of elitism on the natural selection as well as on mutations. This ensures the survival of the best individual so that we do not lose a better solution. This approach is very common in the field of GAs \citep{Haupt2004}. After the natural selection operator, if the best individual has not been selected, it is copied to the mating pool instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower score than the original, the latter is reinserted instead of an individual randomly chosen.

@@

GAs are very computationally intensive because they require many evaluations of the objective function. These assessments are very long in our application, as they require calculating and assessing a forecast for every day of the calibration period. In order to reduce the computation time, we avoid recalculating the score of an individual who has previously been evaluated and that has not changed. We keep the score of each individual living in the selection until it mutates.

@@

\begin{itemize}
	\item mean score: average of the final scores of the 10 optimizations,
	\item convergence: the number of optimizations that converged to a supposed global optimum,
	\item number of generations: characterization of the convergence speed,
	\item number of evaluations of the objective function: characterization of the required calculation time (more realistic than the number of generations).
\end{itemize}

@@

\begin{itemize}
	\item \textbf{Population size}: we found the following ranges to be accurate in average:

@@

	\item \textbf{Selection of couples}: 6 variants of the couples selection were assessed. The performance of these variants are relatively close, both in terms of score, convergence, and number of evaluations. The random pairing performed the most poorly, when the tournament selection with 3 candidates was slightly superior. The roulette wheel weighting is not far behind, but it is less effective in terms of convergence and number of evaluations. This operator has not a significant role in our application.

@@

	The mutation operators based on a variable normal or uniform law work very poorly and are difficult to configure. We then observe many operators more or less with the same scores and requiring a variable amount of assessments. The convergence analysis allows us to highlight three best operators:

@@

The Forecaster and the Viewer are fully available, whereas the open version of the Optimizer includes all methods but the Genetic algorithms. These are planed to be released in the future, but not at the time of writing, due to lacking documentation and necessary return on investment made by Terranum.

>>>>

When using a parameterization of the AM, previously optimized in perfect prognosis \citep[optimization on the reanalysis dataset, without taking into account the uncertainty of operational forecasts,][]{Klein1963}, in operational forecasting, we should take into account the increasing uncertainty of the predictor with the lead time. An option, proposed by \citet{Thevenot2004}, is to increase the number of analogues with the lead time. These values should be optimized, for every lead time, by using reforecast data of the model considered for operational forecasting.

@@

Another reason for working in a perfect prognosis framework is that numerical models evolve continuously, and so does the forecast they provide. Considering re-forecasts would allow us to work on a homogeneous dataset, as they are regularly reprocessed with a defined version of the model. However, we would need to redo the calibration procedure every time a new version is available, in order to reduce the bias \citep{Wilson2002}. Moreover, the reforecasts are not re-processed for every new version of the model, meaning we still end up with a bias between the forecast and the archive. Finally, the length of reanalyses datasets are usually much longer than reforecast datasets, which allows us to identify more robust relationships. The size of the archive is indeed an important criteria for the AM.

@@

We mainly rely on the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows assessing the predicted cumulative distribution functions $F(y)$, for example of the precipitation values from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

\begin{equation}
\label{eq:CRPSS}
CRPSS = \dfrac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\dfrac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference and $CRPS_{p}$ would be the one for a perfect forecast (which implies $CRPS_{p}~=~0$).

@@

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criteria) over a large domain. Every point (or cell) of the full domain is assessed jointly on all predictors of the level of analogy (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The detailed stages are the followings: (i) The unitary spatial window is expanded in every 4 directions successively. The performance score is processed for these 4 windows. (ii) Only the direction providing the best improvement is applied to our spatial window. (iii) From this new spatial window, an increase in every 4 directions is once again assessed, and the best improvement is applied. (iv) The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analogue situations is optimized for the current level of analogy.
		\item A new level of analogy can be added, based on other variables (such as the moisture index) on predefined pressure levels and time frames. The number of analogues for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (a) for the new level. The parameters calibrated on the previous analogue levels are fixed and do not change (except the number of analogues, at the final stage). 
	\end{enumerate}
	\item Finally, the numbers of analogues are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogues of each level in a systematic way.
\end{enumerate}

@@

There are numerous implementation variants of GAs often optimal for a given problem \citep{Hart1991a,Schraudolph1992a}. The divergences are the operators implementation, through significantly different algorithms, which has an important effect on the results \citep{Gaffney2010a}. However, the structure of the method (Figure \ref{fig:structure_gas}) resulting from the work of \citet{Holland1992b} is common to most applications \citep{Back1993b}, and consists in the following steps:

\begin{enumerate}
	\item A population of $N$ individuals (parameter sets of the AM) is randomly generated, which constitutes the initial population.
	\item The fitness (performance score or objective function) of every individual is assessed.
	\item A natural selection is applied, after which only the best individuals remain, which constitutes the intermediate generation.
	\item From this intermediate generation, couples are formed according to given rules.
	\item These couples proceed to reproduction, or chromosome crossover, in order to mix their genes (parameters) according to the operator version. New children are generated in order to fill the intermediate generation back to $N$ individuals.
	\item Parents and children are then subject to mutation, where some genes get affected, meaning some parameter values are randomly changed.
	\item The new formed generation is then re-assessed.
	\item According to the ending criteria, the optimization ends or starts again for another iteration.
\end{enumerate}

@@

	\item \textbf{Tournament selection} \citep{Michalewicz1996, Zitzler2004a}: two individuals are randomly selected and fight. The one with the highest performance score is chosen, but with a certain probability, in order to reduce the selection pressure. This procedure is repeated until the mating pool is full. Individuals can be selected several times, and thus be represented several times in the mating pool.
\end{itemize}

@@

	\textit{Roulette wheel weighting on fitness}: the selection probability is calculated based on the value of the objective function. This approach gives more weight to the best individuals when the distribution of performance scores is wide, while the weight is almost similar when all individuals have approximately the same performance score \citep{Haupt2004}. The probability $p_{n}$ of each individual is calculated by the equation \ref{equation_mating_score_weighting}:
	\begin{equation}
	p_{n}=\frac{score_{n}-score_{N_{mp}}}{\sum_{n=1}^{N_{mp}} (score_{n}-score_{N_{mp}})}
	\label{equation_mating_score_weighting}
	\end{equation}
	In our application, the last individual ($N_{mp}$) has zero probability of being selected.

@@

	\item \textbf{Individual adaptive search radius} (new): based on the ideas of the non-uniform mutation, we introduce a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1, is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is considered initially as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{equation_mutation_rayon_adaptatif}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, in the same way as for the non-uniform mutation. No external parameter is therefore necessary.

@@

	\item \textbf{Chromosome of adaptive search radius} (new): we introduced this operator that combines the operations of the chromosome of adaptive mutation rate to our adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second containing the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameters are required.

	\item \textbf{Multi-scale mutation} (new): finally, we developed another approach, that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea is consistent as long as we are confident that the algorithm will converge towards the global optimum. Indeed, once the algorithm is in the exploitation mode, it is very unlikely to go out of the region it converges to. We wanted to test an approach that combines both exploration and exploitation during the whole optimization. Thus, we considered the search radius $r_{a}$ of equation \ref{equation_mutation_rayon_adaptatif} as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02, which range from full exploration to fine exploitation. The only external parameter is the mutation rate which is fixed.

@@

We used a process of elitism on the natural selection as well as on mutations. This ensures the survival of the best individual so that we do not lose a better solution. This approach is very common in the field of GAs \citep{Haupt2004}. After the natural selection operator, if the best individual has not been selected, it is copied to the mating pool instead of an individual randomly picked. After mutation, if the best individual has mutated and if its new version has a lower performance score than the original, the latter is also reinserted in the mating pool instead of an individual randomly chosen.

@@

GAs are very computationally intensive because they require many evaluations of the objective function. These assessments are very long in our application, as they require calculating and assessing a forecast for every day of the calibration period. In order to reduce the computation time, we avoid recalculating the performance score of an individual who has previously been evaluated and that has not changed. We keep the score of each individual living in the selection until it mutates.

@@

\begin{itemize}
	\item mean performance score: average of the final scores of the 10 optimizations,
	\item convergence: the number of optimizations that converged to a supposed global optimum,
	\item number of generations: characterization of the convergence speed,
	\item number of evaluations of the objective function: characterization of the required calculation time (more realistic than the number of generations).
\end{itemize}

@@

\begin{itemize}
	\item \textbf{Population size $N$}: we found the following ranges to be accurate in average for the population size (number of parameter sets of the AM):

@@

	\item \textbf{Selection of couples}: 6 variants of the couples selection were assessed. The performance of these variants are relatively close, both in terms of performance score, convergence, and number of evaluations. The random pairing performed the most poorly, when the tournament selection with 3 candidates was slightly superior. The roulette wheel weighting is not far behind, but it is less effective in terms of convergence and number of evaluations. This operator has not a significant role in our application.

@@

	The mutation operators based on a variable normal or uniform law work very poorly and are difficult to configure. We then observe many operators more or less with the same performance scores and requiring a variable amount of assessments. The convergence analysis allows us to highlight three best operators:

@@

The Forecaster and the Viewer are fully available, whereas the open version of the Optimizer includes all methods but the Genetic algorithms. These are planed to be released in the future, but not at the time of writing.
